%------------------------------------------------------------------------------
%Description       : DDMoRe WP7.2.1 Functional Specification for the Model
%                                   Repository Infrastructure - Final Points
%Author            : Mihai Glonț <mglont@ebi.ac.uk>
%Organization      : EMBL-EBI
%                    Wellcome Trust Genome Campus
%                    Hinxton
%                    Cambridge
%                    United Kingdom
%------------------------------------------------------------------------------
\section{Final Considerations} 

\subsection{Software Infrastructure}
\label{softwareInfrastructure}
The \ddmore Model Repository stores models under version control on the file system, while the metadata associated with them is maintained in a relational database. Consequently, an instance of the Repository may be deployed along with a database server and a web server communicating over LAN. The models within the Repository should be contained in a centralised location which should be accessible from the web server.

\subsection{Security Considerations}
\label{securityConsiderations}
The security of the data centre, including servers, networking and storage, as well as data, is the sole responsibility of the Administrators that install and maintain an instance of the \ddmore Model Repository. This section is only concerned with the security of the Repository itself. 

The Repository requires write-access to the centralised location managed by a version control system(VCS) where models and associated data will be stored. The VCS tracks the evolution of the content it manages by logging additional metadata in sensitive folders. Administrators must ensure that it is not possible to access this metadata outside of the Repository's interface and, in particular, that the sensitive locations are not available via HTTP. 

Accordingly, the Repository should not need administrative privileges. However, if elevated access rights are required, the functionality that relies upon them should be minimised and the developers should supply Administrators with samples of any special configuration policy that is needed. The Repository will also require superuser privileges on a designated relational database. Access to any other databases or administrative tasks is not needed.

Finally, the Repository will send and receive traffic via HTTP and HTTPS and will communicate with a database server and a messaging queue. If additional ports are needed by an implementation of the \ddmore Model Repository, the developers will need to agree with the Administrators on the choice of port numbers before the implementation goes into production. 

\subsection{Licence}
\label{licence}
The code for the \ddmore Model Repository will be hosted on SourceForge and released under a GPL-compatible licence. This licence does not apply to the models stored in either the public Repository or any private instance of it. 
